---
title: "Final Project Step 3: Data Generalization"
author: "Zeke Krug"
date: "4/25/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load(file="mapDataList.rdata")
load(file="mapAgentData.rdata")
```

## Creating and Using Custom Functions

I've created several custom functions in order to traverse my data.  Currently, the most important 
data is being stored in what R calls a "Large list".  Its a nested list (called mapDataList) which holds all of the data.frames for a given rank on a given map.

```{r}
head(mapDataList, 1)
```
In order to (hopefully) make it clearer the structure of the list, if I wanted to find the data
for the a given map, I would use the following R commands:  
`mapDataList[[i]]` -> a list of 18 data.frames representing the data for a given rank (since
there are 7 maps, i should be in 1:7 and I have a way to make sure of this as well.)  
Then, to get the data for a given rank, I would take the list I got back from `mapDataList[[i]]`
and do:
`map<-mapDataList[[i]]` (this gets the map we're looking for)  
`rank<-map[[j]]` (j should be in 1:18 since I only have data for 18 ranks, and I also have a way to
guarantee this as well.)  

The way that I can guarantee that is by having pre-defined vectors for each map and rank,
and then using the `which` function to be able to find the given index where that map/rank
is in the list (`mapDataList` has been set up so that the maps and ranks are in the same order
as defined in their respective vectors).

```{r}
(maps<-c("ascent", "fracture", "haven", "icebox", "lotus", "pearl", "split"))
(agents<-c("Astra", "Breach", "Brimstone", "Chamber", "Cypher", "Fade", "Gekko",
          "Harbor", "Jett", "KAY/O", "Killjoy", "Neon", "Omen", "Phoenix", "Raze",
          "Reyna", "Sage", "Skye", "Sova", "Viper", "Yoru"))
(ranks<-c("Iron 1", "Iron 2", "Iron 3", "Silver 1", "Silver 2", "Silver 3", "Gold 1", "Gold 2", "Gold 3", "Platinum 1", "Platinum 2", "Platinum 3", "Diamond 1", "Diamond 2", "Diamond 3"))
```



I also have another Large list that holds all of the agent data from each rank and map. 

```{r}
head(mapAgentData, 1)
```
This one is even harder to try to explain, but here goes.  
The first level of lists are the seven maps, so `mapAgentData[[i]]` gets you into a certain map.  
Then if you go one more level down, `mapAgentData[[i]][[j]]`, you get into the agents (of which there are 21).  
And finally, the last level, `mapAgentData[[i]][[j]][[k]]` is where you'll find data for the agent on the specified map for all 18 ranks (or levels of play).
`mapAgentData` has also been set up to be in the same order as the above vectors.  

**NOTE:** *Actually, I just realized how terrible this is and a point for optimization.  I'm going to convert each list at the last level into a data.frame so there's only two levels that need to be traversed.  So `mapAgentData[[i]][[j]]` will be a data.frame for the given character containing the data from all ranks in the order specified by the `ranks` vector. Every entry in the original `mapAgentData[[i]][[j]][[k]]` will become a row in the data.frame.*

# Some functions from Step 2

These are some functions that might be useful in producing interesting stats to the user

```{r}
# agent (string, name of agent) -> integer (the number of matches the given agent played on ascent)
# This is a map specific function, however, I believe I can generalize it more
# to take a map and an agent.
getNumberOfMatchesAscent<-function(agent){
  selectedAgent<-which(agents==agent)
  agentData<-ascentAgentData[[selectedAgent]]
  matches<-list()
  for(i in 1:length(agentData)){
    element<-agentData[[i]]
    matches[i]=element$Matches
  }
  return(sum(unlist(matches)))
}


#   -> list containing the number of times each agent had the highest win % on ascent
# This is another map specific function that can be changed to 
# be more generalized to take an agent and map
getNumberOfHigestWinPrcntAscent<-function(){
  agentList<-list()
  mapData<-mapAgentData[[1]]
  for(i in 1:length(agents)){
    currentAgentData<-mapData[[i]]
    countTopRank<-0
    for(j in 1:length(currentAgentData)){
      element<-currentAgentData[[j]]
      if(element$Rank==1){
        countTopRank<-countTopRank+1
      }
    }
    agentList[i]=countTopRank
  }
  return(agentList)
}

# The main function that I want to implement in order to display to a user
# rank (string, rank which represents the level of play), map (name of map) -> data.frame of the top 5 agents for the map at the given rank
processRequest<-function(rank, map){
  selectedMap <- mapDataList[[which(maps == map)]]
  selectedRank <- selectedMap[[which(ranks == rank)]]
  result <- getTop5(selectedRank)
  return(result)
}

# helper function for the main function to get rid of unimportant data
# df (data.frame representing the data for a given rank and map) -> subsetTop5 (subset of original data.frame containing the first 5 rows and specified columns)
getTop5<-function(df) {
  top5<-head(df, 5)
  subsetTop5<-subset(top5, select=c("Agent", "Win..", "Avg..Score","KD", "Matches"))
  colnames(subsetTop5)[2] = "Win Prct"
  colnames(subsetTop5)[3] = "Average Combat Score"
  return(subsetTop5)
}

```




## Shiny App Overview

For the Shiny application, I plan to convert some of my predefined functions that I made
for Step 2 and make them take a few extra parameters like map and rank in order to output the
team with the best chance to win and maybe a few extra statistics (if I have time).  

# UI
For the UI, I plan to have two inputs, one for rank and one for map.  I'm not sure what kind of input
makes sense yet, probably not a slider though, but I'll figure it out with some trial and error.

# Server
For the server, I want to have a couple of functions that do different things to display specific statistics to the user, 
but the main function (`processRequest` above) will just present the agents with the highest win % for the given map and rank.

